A generic and fast C++ optimization framework

### 2017-10-19 paper.tex
Refactor table, add a few more parts to it. Tilt the text a bit less for easy reading, use \adjustbox to make the width right.

\newcommand*\rot{\rotatebox{90}}
\newcommand*\OK{\ding{51}}

@@

\author{
  Shikhar Bhardwaj \\
  \AND
  Yannis Mentekidis \\
  \AND
  Marcus Edel \\
  Free University of Berlin \\
  Arnimallee 7, 14195 Berlin \\
  \texttt{marcus.edel@fu-berlin.de} \\
  \AND
  Ryan R. Curtin \\
  Symantec Corporation \\
}

@@

However, the current landscape of optimization frameworks for machine learning
is not particularly comprehensive.  A variety of tools such as
Caffe \cite{jia2014caffe},
% mxnet ??
TensorFlow \cite{abadi2016tensorflow},
and
Keras \cite{chollet2015}
have optimization frameworks, but they are limited to SGD-type optimizers and
are only able to optimize deep neural networks or related structures.  Thus
expressing arbitrary machine learning objective functions can be difficult or in
some cases not possible.  Other libraries, like
% Shogun \cite{shogun}
%and
scikit-learn \cite{pedregosa2011scikit},
do have optimizers implementation, but generally not in a coherent framework and
often the implementations may be specific to an individual machine learning
algorithm.  At a higher level, many programming languages may have generic
optimizers, like SciPy \cite{jones2014scipy} and MATLAB, but typically these
optimizers are not suitable for large-scale machine learning tasks where, e.g.,
calculating the full gradient of all of the data may not be feasible.

@@

% Just to get a feeling of the look; Also, we could also just use \OK instead of \CURCLE or \LEFTcircle
\begin{table}
\centering
    \begin{tabular}{@{} cl*{6}c @{}}
        & & \multicolumn{6}{c}{} \\[2ex]
        & & \rot{has framework}
          & \rot{\shortstack[l]{problem-independent\\constraints}}
          & \rot{batches}
          & \rot{reusability}
          & \rot{x}
          & \rot{y} \\
        \cmidrule{2-8}
        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE  \\
        & vowpal wabbit      & - & - & \LEFTcircle  & \CIRCLE & - & - \\
        & TensorFlow         & - & -  & \CIRCLE  & \LEFTcircle & - & -  \\
        & scikit-learn       & - & -  & \CIRCLE  & \LEFTcircle & - & - \\
        & scipy              & - & \LEFTcircle  & -  & \LEFTcircle & - & - \\
        \cmidrule[1pt]{2-8}
    \end{tabular}
    \begin{tablenotes}\footnotesize
\item[*] Library feature comparison: \CIRCLE=\text{provides feature}, \LEFTcircle=\text{partially provides feature}, -=\text{does not provide feature}
\end{tablenotes}
\end{table}

>>>>

\usepackage{adjustbox}
\usepackage{array}

\newcommand*\OK{\ding{51}}

\newcolumntype{R}[2]{%
  >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
  l%
  <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{60}{2em}}}

@@

\author{
  Shikhar Bhardwaj \\
  \AND
  Yannis Mentekidis \\
  \AND
  Marcus Edel \\
  Free University of Berlin \\
  Arnimallee 7, 14195 Berlin \\
  \texttt{marcus.edel@fu-berlin.de} \\
  \AND
  Ryan R. Curtin \\
  Symantec Corporation \\
  \texttt{ryan@ratml.org} \\
}

@@

However, the current landscape of optimization frameworks for machine learning
is not particularly comprehensive.  A variety of tools such as
Caffe \cite{jia2014caffe},
% mxnet ??
TensorFlow \cite{abadi2016tensorflow},
and
Keras \cite{chollet2015}
have optimization frameworks, but they are limited to SGD-type optimizers and
are only able to optimize deep neural networks or related structures.  Thus
expressing arbitrary machine learning objective functions can be difficult or in
some cases not possible.  Other libraries, like
% Shogun \cite{shogun}
%and
scikit-learn \cite{pedregosa2011scikit},
do have optimizers implementation, but generally not in a coherent framework and
often the implementations may be specific to an individual machine learning
algorithm.  At a higher level, many programming languages may have generic
optimizers, like SciPy \cite{jones2014scipy} and MATLAB, but typically these
optimizers are not suitable for large-scale machine learning tasks where, e.g.,
calculating the full gradient of all of the data may not be feasible.  Table
\ref{tab:features} provides a rough overview of the capabilities of many common
machine learning libraries and languages.

@@

% Just to get a feeling of the look; Also, we could also just use \OK instead of \CURCLE or \LEFTcircle
\begin{table}
\centering
    \begin{tabular}{@{} cl*{7}c @{}}
%  \begin{tabular}{ccccccc}
        & & \multicolumn{7}{c}{} \\[2ex]
            % If there is any coherent framework at all, this is true.
        & & \rot{has framework}
            % If there is any support for constrained optimization, this is
            % true.
          & \rot{constraints}
            % If the optimization framework can do mini-batch, this is true.
          & \rot{batches}
            % If I can implement any arbitrary function to be optimized, this is
            % true.
          & \rot{arbitrary functions}
            % If I can implement any new optimization technique to use, this is
            % true.
          & \rot{arbitrary optimizers}
            % If the framework could take advantage of when the gradient is
            % sparse, this is true.
          & \rot{sparse gradients}
            % If the framework can handle categorical/discrete variables for
            % optimization, this is true.
          & \rot{categorical} \\
        \cmidrule{2-9}
  \hline
        % It might be reasonable to say mlpack categorical support is only
        % partial, but I am not sure exactly where we draw the line.
        & mlpack             & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE & \CIRCLE \\
        % The Shogun toolbox has a fairly nice framework, but it doesn't support
        % sparse gradients or categorical features.  It also does not appear to
        % support constraints.
        & shogun             & \CIRCLE & - & \CIRCLE & \CIRCLE & \CIRCLE & - & -
\\
        % VW doesn't appear to have any framework whatsoever and the code is
        % awful, but it does support batches and categorical features.
        & vowpal wabbit      & - & - & \CIRCLE  & - & - & - &
\CIRCLE \\
        % TensorFlow has a few optimizers, but they are all SGD-related.  You
        % can write most objectives easily (but some very hard), and categorical
        % support might be possible but would not be easy.
        & TensorFlow         & \CIRCLE & -  & \CIRCLE  & \LEFTcircle & - &
\LEFTcircle & -  \\
        % Caffe has a nice framework, but it's only for SGD-related optimizers.
        % I think I could write a new one, but it is not the easiest thing in
        % the world.
        & caffe             & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
& - & - \\
        % Keras is restricted to neural networks and SGD-like optimizers.  I
        % don't know that it is possible to easily write a new optimizer.
        & keras             & \CIRCLE & -  & \CIRCLE & \LEFTcircle & \LEFTcircle
& - & - \\
        % sklearn has a few optimizer frameworks, but they are all in different
        % places and have somewhat different support.
        & scikit-learn       & \LEFTcircle & - & \LEFTcircle  & \LEFTcircle & -
& - & - \\
        % scipy has some nice optimizer framework but it does not support
        % batches or some of the more complex functionality.  And you can't
        % write your own.
        & scipy              & \CIRCLE & \CIRCLE  & -  & \CIRCLE & - & - & - \\
        % MATLAB is very similar to scipy.
        & MATLAB             & \CIRCLE & \CIRCLE & - & \CIRCLE & - & - & - \\
        \cmidrule[1pt]{2-9}
    \end{tabular}
%   \begin{tablenotes}\footnotesize
\caption{
Library feature comparison: \CIRCLE = provides feature,
\LEFTcircle = partially provides feature, - = does not provide feature.
{\it has framework} indicates that the library has some kind of generic
optimization framework; {\it constraints} and {\it batches} indicate support for
batches; {\it arbitrary functions} means arbitrary objective functions are
easily implemented; {\it arbitrary optimizers} means arbitrary optimizers are
easily implemented; {\it sparse gradient} indicates whether the framework can
natively take advantage of sparse gradients; and {\it categorical} refers to
whether support for categorical features exists.
}
\label{tab:features}
\vspace*{-1.5em}
\end{table}
